\vspace{3cm} \Huge{\futhark}\\ \Large{Enhancing performance by rewriting kernels from Python/OpenCL to Futhark}

### 2017-11-09 report/draft.tex
upw

\section{Demonstration of use of finished project}
As \sasmodels is rarely used on its own, but rather as the backbone of 
\sasview's calculation backend, I am not going to write much about 
\sasview usage.
Instead, I will describe how Futhark models in \sasmodels are leveraged 
by \sasmodels' users.
måske ikke så meget mere her
% mere

\subsection{Describing \sasmodels' models}
A \sasmodels model info file (hereafter referred to as just a model info) is a 
Python file, that defines at least two primary properties of a 
calculation model \textit{m}:
\begin{itemize}

  \item a function definition \texttt{I(q)}, which defines scattered intensity 
  of the scattering vector \texttt{q}.

@@

\begin{figure}
  % fuld beskrivelse af ax+b-figur, med parametre a, b, og en Iqxy
  \caption{model info file for modelling a line}\label{fig:linemodel}
\end{figure}

@@

\subsection{Using a Futhark model}
\label{sec:using-futhark}
There is a difference between defining and using a Futhark model.
The \sasview user who wishes to use a Futhark model in his or her work does 
not need more than the futhark model info file, and a copy of the compiled
Futhark kernel. In fact, the futhark model info file is identical to OpenCL- 
and Python model info files, except for in two properties.
As show in figure \ref{fig:futhark_dab}, there has been added a flag 
%CHECK THISSS!!!
\texttt{model\_is\_futhark: true}, and \iq is now a path to a precompiled 
Futhark kernel.

The user does not need to have any Futhark compiler installed if he merely wants
to use the precompiled Futhark module: Futhark models for \sasmodels compiles to
portable Python libraries which has numpy and pyopencl as their only 
dependencies - and as \sasmodels has these dependencies as well, the user will
already have these dependencies sorted out, if he has a working installation of 
\sasview or \sasmodels.

@@

Other models, such as core-shell %FIX\
imports libraries so they are available in the Iq function.
%% som vist i core shell, kan det også være nødvendigt at definere og importere 
% ekstra biblioteker

@@

\section{Implementation of \texttt{kernelfut.py}}
\sasmodels

% current architecture of Sasmodels engine
% description of kernelcl kernels
% (smaller) description of kernelcl kernels
% description of current kernelfut implementation

@@

% nævn at projektet er blevet præsenteret for en faktisk brugergruppe,
% og fik god og konstruktiv feedback
% det inspirerede til yderligere projekter/forbedringer indenfor sasmodels

@@

>>>>

\section{Demonstration}
\begin{mdframed}[frametitle={Terminology:}]
  The work of this project has been spread out across multiple parts
  of the \sasmodels program. Here is a list of the distinct parts of \sasmodels
  that has been worked with:
  \begin{itemize}
    \item \texttt{kernelfut}: The \sasmodels Futhark kernel module that wraps 
    \sasmodels input data and sends it to be calculated in the actual Futhark 
    kernel. \texttt{kernelfut} is an alternative to \texttt{kernelpy} and
    \texttt{kernelcl} which are already available in \sasmodels.

    \item Model info files: Every model in \sasmodels is defined in a Python 
    model info file, which defines all the model's properties, like the model's
    function itself, and a list of the parameters needed to run the model.
    See figure \ref{fig:dabmodel}

    \item Futhark models: A Futhark model is the model module that contains the
    functions needed to run a given model. See the Futhark line model in fig. 
    \ref{fig:fig:linemodel-futhark}

    \item The Futhark kernel template, as shown in figure \ref{fig:for-iq}:
    The Futhark kernel template contains the module that takes a Futhark
    model as input, and returns a Futhark kernel that can be called by 
    \texttt{kernelfut}.

    \item Futhark kernels: The Python OpenCL library that is produced by 
    compiling the Futhark kernel template with a Futhark model.

    \item \sasmodels experiments: A \sasmodels experiment is the action of 
    running \sasmodels with some input data and one or more models, to get 
    the resulting output data. The output data is the selected models applied
    to the input data.
  \end{itemize}

\end{mdframed}

\subsection{Describing \sasmodels' models}
\begin{itemize}
  \item a function definition \texttt{I(q)}, which defines scattered intensity 
  of the scattering vector \texttt{q}.

@@

@@

\subsection{Using a Futhark model}
\label{sec:using-futhark}
There is a difference between defining and using a Futhark model.
The \sasview user who wishes to use a Futhark model in his or her work does 
not need more than the futhark model info file, and a copy of the compiled
Futhark kernel. In fact, the futhark model info file is identical to OpenCL- 
and Python model info files, except for in two properties.
As show in figure \ref{fig:linemodelinfo-futhark}, \iq is now a dictionary
with the single flag \texttt{model\_is\_futhark} set to \texttt{True},
and the property \texttt{futhark\_path} points to the location of the compiled
Futhark kernel.

The user does not need to have any Futhark compiler installed if he merely wants
to use the Futhark kernel: \futhark compiles to portable Python libraries which 
has numpy and pyopencl as their only dependencies - and as \sasmodels has these 
dependencies as well, the user will already have these dependencies sorted out, 
if he has a working installation of \sasview or \sasmodels.

@@

Other models, such as core-shell imports libraries so they are 
available in the Iq function.

@@

\section{Adding \texttt{kernelfut} to \sasmodels}
The following process happens when running a \sasmodels{} experiment through the
built-in comparison module:
\begin{enumerate}
  \item The experiment parameters\footnote{choice(s) of model, number of 
  evaluations, sample size, number type and so on} for the experiment are used
  to create a ParameterTable.

  \item The ParameterTable is used to load (or generate) a set of input data for
   the kernel.

  \item The ParameterTable is used to initialize a \sasmodels kernel using the
  desired model. In case of OpenCL and Futhark kernels, the input data is 
  buffered to the OpenCL context already at this point. This is done to leave
  as much buffering out of the time measurements of \sasmodels as possible.

  \item \sasmodels packs the model and the data in a DirectModel wrapper object.

  \item The DirectModel is run the desired number of times in \sasmodels,
  and finally returns the model output, the average running time, and the
  output difference between the models that have been compared. 
\end{enumerate}

To design \texttt{kernelfut}, it was first necessary to ape the 
\sasmodels calculation loop as a Futhark function. 
I chose to replicate \texttt{kernelpy}'s \texttt{\_loops(...)} function in 
Futhark, as it is fairly simple to rewrite the Python code in Futhark.
\texttt{\_loops()} is shown in figure \ref{fig:kernelpy_loop}, and the
comparable Futhark function in figure \ref{fig:kernelfut_loop}.

Second, it was necessary to implement a bridge between \sasmodels and the 
precompiled Futhark kernels. 
When a \sasmodels experiment wants to use a Futhark kernel for it's calculation, it must dynamically import the kernel
% gør når jeg kan koncentrere mig

% (smaller) description of kernelcl kernels
% description of current kernelfut implementation

@@

@@

\begin{figure}
  \lstinputlisting[language=Python, linerange={180-252}]{../sasmodels/sasmodels/kernelpy.py}
  \label{fig:kernelpy_loop}
  \caption{The Python kernel calculation loop for \sasmodels}
\end{figure}

\begin{figure}
  \lstinputlisting[linerange={13-53}]{../sasmodels/sasmodels/models/futharks/header/for_iq.fut}
  \label{fig:kernelfut_loop}
  \caption{The Futhark kernel calculation loop, aping fig \ref{fig:kernelpy_loop}}
\end{figure}
